# Week 2: Regression and model validation

## Reading the students2014 data into R    
### R-code and output   
    
You can find the code pressing the "Code"-button on the right ->
```{r}
getwd()
setwd("/Users/mirva/IODS-project/data")
learning2014 = read.csv("learning2014")
dim(learning2014)
str(learning2014)
```
    
This data is part of the data which Kimmo Vehkalahti has collected 3.12.2014 - 10.1.2015 on the course:
    
Johdatus yhteiskuntatilastotieteeseen, syksy 2014 (Introduction to Social Statistics, fall 2014 - in Finnish)     
     
The data has 166 observations and 7 variables which are gender, age, attitude, deep, stra, surf and points    

* gender has two options: M (Male), F (Female)    
* age is (in years) derived from the date of birth    
* attitude describes global attitude toward statistics    
* deep, stra and suf are on the Likert-scale    
* deep refers to deep learning and is the average score of the subsets Seeking Meaning, Relating Ideas and Use of Evidence     
* stra refers to strategic learning and is the average score of the subsets Organized Studying and Time Management  * surf refers to surface learning and is the average of the subsets Lack of Purpose, Unrelated Memorising and Syllabus-boundness     
* points are the exam points of the course    
    
    
***
    
    
## Graphical overview of the data
### R-code and plot
    
```{r}
library(ggplot2)
library(GGally)
overw <- ggpairs(learning2014, mapping = aes(col = gender, alpha = 0.3), 
             lower = list(combo = wrap("facethist", bins = 20)))
overw
summary(learning2014)
```

    
The variables are all quite normally distributed except the age-variable where the min is 17, median 22 and max 55. Also the differences between men and women are quite small, allthough the distributions are not exactly equal. Biggest overall correlations are between attitude and points (r=.437) and deep and surf (negative, r=-.324), correspondingly smallest overall correlations are between deep and points (negative, r=-.0101), attitude and age (r=.0222) and deep and age (r.0251). 
    
    
***
    
    
## Choosing the variables and fitting a regression model    
    
y="points", x1="attitude", x2="stra" and x3="surf"    
    
    
### Model 1    

```{r}
ggpairs(learning2014, lower = list(combo = wrap("facethist", bins = 20)))
regr1 <- lm(points ~ attitude + stra +surf, data = learning2014)
summary(regr1)
```
    
T-value of attitude (5.913) is statistically significant (p<.001) so the attitude-variable explains well the points-variable. T-values of stra (1.575) and surf(-0.731) aren´t significant in this model so I will remove the one with smaller t-value: the variable surf.
    
    
### Model 2    

```{r}
regr2 <- lm(points ~ attitude + stra, data = learning2014)
summary(regr2)
```
    
The p-value of the variable stra (p=.08927) remains still >.05 so i will also remove it from the model.
    
    
### Model 3 
    
```{r}
regr3 <- lm(points ~ attitude, data = learning2014)
summary(regr3)
```
    
    
***
    
    
## Interpretation of the model 
    
The relationship between attitude and points: when y = "attitude" and x = "points" we can predict the value of y with model: 
    
y = 11.63715 + 0.35255x + e    
    
Multiple R-squared of this model is 0.1906, so the model explains 19,06% of the variation of attitude
    
    
***
    
    
## Diagnostic plots
    
```{r}
par(mfrow = c(2,2))
plot(regr3, which = c(1,2,5))
```
   
    
### Interpretation of the plots
* Residuals vs Fitted: Residuals are quite randomly around the 0 line so it could indicate that the residuals and fitted values are not correlated.
* Normal QQ-plot: At most part the Q-Q plots follow the theoretical line quite nicely and are therefore normally distributed, but at the beginning and in the end of the line the plots aren´t so well on the line which indicates that there is light left skewness
* Residuals vs Leverage: There seems to be three outliers (35, 56 and 71) which could influence the model    
    
    
    
    