# Week 3: Logistic regression
```{r}
library(ggplot2); library(dplyr); library(GGally); library(boot)
```
    
    
## Reading the data into R    
    
At first I´ll read the alcohol data into R
```{r}
getwd()
setwd("/Users/mirva/IODS-project/data")
alc = read.csv("alcohol")
```
    
### The structure of the data
```{r}
str(alc)
```
     
This data includes 382 observations and 36 variables    
The data are from two identical questionaires related to secondary school student alcohol comsumption in Portugal. 
    
### Database info
    
Using Data Mining To Predict Secondary School Student Alcohol Consumption.  
Fabio Pagnotta, Hossain Mohammad Amran.    
Department of Computer Science,University of Camerino    
You can find more info about the variables and data [here](https://archive.ics.uci.edu/ml/datasets/STUDENT+ALCOHOL+CONSUMPTION)
    
    
## The hypotheses
    
I will study the relationship between high/low alcohol consumption and four other variables: Pstatus, Medu, Fedu and famrel    
    
My personal hypothesis about these four variables and alcohol consumption are as follows:    
Pstatus = parents´ cohabitation status (living together or apart): My hypothesis is that students whose parents are living together could have lower risk of high alcohol consumption.
Medu and Fedu = mother's education and father´s education: My hypothesis is that parents´lower education (and therefore possibly lower socio economical status) could increase the risk of high alcohol consumption.    
famrel = quality of family relationships: My hypothesis here is that students whose family relationships are graded 4 or 5 could have lower risk of high alcohol consumption.
    
    
## Exploring the distributions
### Parents´cohabilitation status
```{r}
alc %>% group_by(Pstatus, high_use) %>% summarise(count = n())
```
    
Within students whose parents´cohabilitation status is A = "apart" 26 (68%) students´ alcohol usage is low and 12 (32%) it is high. Within students whose parents live together (status = T) 242 (70%) students´ alcohol usage is low and 102 (30%) it is high.
    
### Mother´s education level
```{r}
alc %>% group_by(Medu, high_use) %>% summarise(count = n())
g1me <- ggplot(alc, aes(x = high_use, y = Medu))
g1me + geom_boxplot() + xlab("High alcohol consumption") + ylab("Mothers education")
```
    
If we look at the mothers education level the results of summary are as follows:

* No education: low drinking n=1 (33%) and high drinking n=2 (67%)
* Primary education (4th grade): low drinking n=33 (65%) and high drinking n= 18 (35%)
* 5th to 9th grade: low drinking n=80 (82%) and high drinking n=18 (18%)
* Secondary education: low drinking n=59 (62%) and high drinking n=36 (38%)
* Higher education: low drinking n=95 (70%) and high drinking n=40 (30%)
    
So there might be some differences between these groups but ns are quite low, especially in non-educated-group.    
The boxplots of these two alcohol consumption groups look exactly the same.    
    
### Father´s education level
```{r}
alc %>% group_by(Fedu, high_use) %>% summarise(count = n())
g1fe <- ggplot(alc, aes(x = high_use, y = Fedu))
g1fe + geom_boxplot() + xlab("High alcohol consumption") + ylab("Fathers education")
```
    
With fathers´education the same numbers are as follows:   

* no education: low drinking n=2 (100%) and high drinking n=0    
* primary education (4th grade): low drinking n=53 (69%) and high drinking n= 24 (31%)    
* 5th to 9th grade: low drinking n=75 (71%) and high drinking n=30 (29%)    
* secondary education: low drinking n=72 (73%) and high drinking n=27 (27%)    
* higher education: low drinking n=66 (67%) and high drinking n=33 (33%)

So also here we can see that fathers´higher education doesn´t possibly imply lower drinking percentages as opposed to my hypothesis.    
    
### Quality of family relations
```{r}
alc %>% group_by(famrel, high_use) %>% summarise(count = n())
g1fr <- ggplot(alc, aes(x = high_use, y = Fedu))
g1fr + geom_boxplot() + xlab("High alcohol consumption") + ylab("Quality of family relations")
```
    
If we look at the connection between alcohol consumption and quality of family relations distribution in different grade groups are:    

* grade 1: low drinkers 6 (75%), high drinkers 2 (25%)    
* grade 2: low drinkers 10 (53%), high drinkers 9 (47%)    
* grade 3: low drinkers 39 (61%), high drinkers 25 (39%)    
* grade 4: low drinkers 135 (71%), high drinkers 54 (29%)    
* grade 5: low drinkers 78 (76%), high drinkers 24 (24%)    

So there might be some support for my hypothesis that better the family relations lower the alcohol consumption    
    
    
## Logistic regression and the relationship between the chosen variables and high/low alcohol consumption
```{r}
m <- glm(high_use ~ Pstatus + Medu + Fedu + famrel, data = alc, family = "binomial")
summary.glm(m)
coef(m)
```
    
Only the variable quality of family relations is significant in this model.
    
### Odds ratios and confidence intervals
```{r}
OR <- coef(m) %>% exp
CI <- confint(m) %>% exp
cbind(OR, CI)
```
    
Every one of these confidence intervals includes 1 except famrel so it implies these other variables make no difference in the model.    
    
Removing non-significant variables one by one doesn´t give any new variables to the model
```{r}
m2 <- glm(high_use ~ Medu + Fedu + famrel, data = alc, family = "binomial")
summary.glm(m2)

m3 <- glm(high_use ~ Fedu + famrel, data = alc, family = "binomial")
summary.glm(m3)
```
    
So I´ll leave just the variable famrel to the model
```{r}
m4 <- glm(high_use ~ famrel, data = alc, family = "binomial")
summary.glm(m4)
```
   
   
## The predictive power of my model
```{r}
probabilities <- predict(m4, type = "response")
alc <- mutate(alc, probability = probabilities)
alc <- mutate(alc, prediction = probability > 0.5)
table(high_use = alc$high_use, prediction = alc$prediction)
g <- ggplot(alc, aes(x = probability, y = high_use, col = prediction))
g2 <- g + geom_point()
g2
table(high_use = alc$high_use, prediction = alc$prediction) %>% prop.table() %>% addmargins()
m <- glm(high_use ~ failures + absences + sex, data = alc, family = "binomial")
```
     
So this means that the predictive power of my model is low (and probably more significant variables should be found to make it more predictive.) As long as all the predictions are under 0.5 the level of prediction is same with guessing.  
    
### Mean prediction error
```{r}
loss_func <- function(class, prob) {
  n_wrong <- abs(class - prob) > 0.5
  mean(n_wrong)
}
loss_func(class = alc$high_use, prob = alc$probability)
```
    
Mean prediction error is approx. 0.298 which means that so many of predictions are false.     
    
    
## 10-fold cross validation
```{r}
cv <- cv.glm(data = alc, cost = loss_func, glmfit = m4, K = 10)
cv$delta[1]
```
    
The average number of wrong predictions in the cross validation is approx. 0.304
    
### An example of a better model
     
Another model where prediction error is smaller compared to the model introduced in DataCamp has variables: absences, sex, famrel and goout. It´s prediction error is about 0.207    
```{r}
m5 <- glm(high_use ~ absences + sex + famrel + goout, data = alc, family = "binomial")
summary.glm(m5)
cv <- cv.glm(data = alc, cost = loss_func, glmfit = m5, K = 10)
cv$delta[1]
```